Binary companions of evolved stars in \apogee\ \DR: \\ Search method and catalog of $\sim$5,000 companions

### 2018-03-20 paper/1-catalog/twoface-catalog.tex
change cut in lnK
Edits and cleaning

\newcommand{\nprior}{536,870,912}
\newcommand{\nposterior}{256}
\newcommand{\nstars}{96,231}
\newcommand{\nvisits}{397,559}
\newcommand{\ncontrol}{16,384}
\newcommand{\nhighK}{4,510}
\newcommand{\nbimodal}{113}
\newcommand{\nunimodal}{319}

@@

\title{Binary companions of red giant stars in \apogee\ \DR: \\
       Search method and catalog of $>$4,500 systems}

@@

\begin{abstract}\noindent % trust me
% Context
Multi-epoch radial velocity measurements of stars can be used to identify
stellar, sub-stellar, and planetary-mass companions.
Even a small number of observation epochs can be useful for detecting
companions, though there can be multiple qualitatively different orbital
solutions that fit the data.
For these reasons, we custom-built a Monte Carlo sampler (\thejoker) that
delivers reliable (and often highly multi-modal) posterior samplings for
companion orbital parameters given sparse radial-velocity data.
% Aims
Here we perform a search for companions to \nstars\ red-giant stars observed in
the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
% Methods
We select stars with probable companions by making a cut on our posterior belief
about the amplitude of the stellar radial-velocity variation induced by the
orbit.
% Results
We provide (1)~a catalog of \nunimodal\ companions for which the stellar
companion properties can be confidently determined, (2)~a catalog of \nhighK\
stars that likely have companions, but for which more observations would be
needed to uniquely determine the orbital properties, and (3)~posterior samplings
for the full orbital parameters for all stars in the parent sample.
We highlight interesting systems and show the characteristics of stars with
confidently determined companion properties.
\end{abstract}

@@

Most of what is known about stellar companions comes from studies of
nearby main-sequence (MS) stars.
\todo{occurrence rates}
MS stars with companions have a large dynamic range of constituent and orbital
characteristics.
For example, binary stars have mass-ratios that span from $\approx 0.03$
to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
millions of years (e.g., \citealt{Raghavan:2010}).
Less is known about population properties of non-interacting or detached
companions to evolved stars.

@@

We have solved this problem previously with \thejoker\
(\citealt{Price-Whelan:2017}), a Monte Carlo rejection sampler that is
computationally expensive but probabilistically righteous:
It delivers independent posterior \pdf samples for single-companion binary
orbital parameters, given any number of radial-velocity measurements.
Here we use \thejoker\ to generate posterior \pdf samples for stars observed by
the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).

The \apogee\ surveys primarily target red-giant-branch (\RGB) stars, which are
ideal for the study of single-line binary systems.
In general, they are unlikely to have equally-bright companions, and their
spectra are therefore fit as single-line objects.
When this constraint is not met, \thejoker\ will in general fail, and a model
that fits for a mixture of stellar spectra is more appropriate (e.g.,
\citealt{El-Badry:2018}).
The subset of giant stars in the ``red clump'' (\RC) are even more powerful as
they are standard candles and have masses that can be estimated using
spectroscopy (using dredged-up elements; \citealt{Martig:2016,Ness:2016}).
With mass estimates for the primary star, the binary-orbit fitting will return
$m_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
of the so-called ``binary mass function.''
Additionally, the \apogee\ pipelines (\citealt{Garcia-Perez:2016}) and also
\thecannon\ (\citealt{Ness:2015}) produce detailed abundance estimates for \RGB\
and \RC\ stars.
If there are causal relationships between chemical abundances and binary
companions---as are expected---these should be measurable.

By making cuts on this library of posterior \pdf samples (described in detail in
\sectionname~\ref{sec:whatever}), we deliver a catalog of binary-star systems
from the \apogee\ survey \todo{with ... (tease some statistics)}.
We additionally highlight the subset of this catalog with \RC\ star primaries,
for which we can compute companion mass and pericentric distance.

@@

All data used in this work comes from the publicly-available data release 14
(\DR) of the \apogee\ survey (\citealt{Majewski:2017,Abolfathi:2017}), a
component of the Sloan Digital Sky Survey IV (\sdssiv;
\citealt{Gunn:2006,Blanton:2017}).
\apogee\ is designed to map stars across much of the Milky Way by obtaining
high-resolution ($R \sim 22,500$) infrared ($H$-band) spectroscopy of primarily
\RGB\ stars.
Targets are selected with simple color and brightness cuts, but the survey uses
fiber-plugged plates with a maximum of 300 fibers per each $\approx
1.5~\textrm{deg}^2$ field of view, leading to ``pencil-beam''-like sampling of
the stellar distribution.
In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
observed multiple times in a series of ``visits,'' typically with at least one
visit separated by a month or more in order to help identify binary stars.

@@

\thejoker\ is a custom-built Monte Carlo sampler designed to produce independent
posterior samples in Keplerian orbital parameters, given radial-velocity
measurements under the assumptions listed above.
Our parametrization of the orbital elements is similar to the notation in
\citet{Murray:2010}:
The radial velocity $v$ at time $t$ is given by
\begin{equation}
  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, M_0, t_0)\right) +
    e\,\cos\omega]
\end{equation}
where $\bs{\theta} = (P,e,M_0,\omega,K,v_0)$---period, eccentricity, mean
anomaly at a reference time $t_0$, argument of pericenter, velocity
semi-amplitude, barycentric velocity---and the true anomaly, $f$, is a function
of time and the specified parameters (see \sectionname~2 of
\citealt{Price-Whelan:2017} or \eqname~63 in \citealt{Murray:2010}).
In addition to the orbital parameters listed above, \thejoker\ can also generate
samples in an ``excess variance'' parameter, $s^2$, that is added to the
per-visit measurement variances.
This parameter allows us to test whether the visit RV uncertainties are
underestimated
For upper \RGB\ stars the inferred excess variance will be a combination of
extra systematic uncertainty and true astrophysical surface jitter
(\todo{CITATION NEEDED}).

\thejoker\ was designed for the extremely multi-modal \pdf s expected when the
number of radial-velocity measurements of a source is small, or the data are
sparse (in phase-coverage) or noisy.
While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
independent samples with such data, \thejoker\ succeeds by brute force:
After generating an initial (very large) library of prior samples from an
assumed prior \pdf (see below), the (typically multi-modal) likelihood is
evaluated at each sample and used to rejection sample.
In practice, given a number of requested samples for each star, the sampling
proceeds iteratively: since it is easier to accept samples when the data is
sparse or noisy, far more prior sample draws (and thus likelihood evaluations)
must occur under very constraining data.

@@

For all \nstars\ \apogee\ stars with $\geq 3$ good visits (see
\sectionname~\ref{sec:data}), we use \thejoker\ to generate posterior samplings
for each star under the assumptions listed above (see
\sectionname~\ref{sec:fitting}).
We start by generating a library of \nprior\ prior samples generated under a
prior similar to that defined in \citet{Price-Whelan:2017}:
\begin{itemize}
    \item uniform or isotropic in angle parameters,
    \item uniform in log-period over the domain $[1,32768]~\textrm{day}$,
    \item a beta distribution over eccentricity (using parameters from
    \citealt{Kipping:2013}).
\end{itemize}
For the excess variance parameter, $s^2$, we initially use a Gaussian over the
transformed parameter $y = \ln s^2$ with the mean and standard deviation
$(\mu_y, \sigma_y)$ indicated where the runs are described (see
\sectionname~\ref{sec:catalogs}).
The reference time for each star is set to the minimum visit epoch; $M_0$ then
becomes the mean anomaly at the first visit observation.
\tablename~\ref{tbl:params} contains descriptions of all parameters and priors
used.

@@

We request \nposterior\ posterior samples for each source.
Depending on the data quality and phase coverage of the visits, \thejoker\ will
require different numbers of prior samples in order to rejection sample down to
the requested number of posterior samples: For few-epoch or noisy RV data, many
prior samples will pass the rejection step, whereas for very precise or
many-epoch RV data, \thejoker\ may need to go process the full library of prior
samples.
We therefore generate the posterior samples using an iterative procedure that
adaptively predicts how many prior samples to test for each star.
For sources with very constraining data, \thejoker\ may return fewer than the
requested number of samples (as few as one sample).
When this occurs, we continue sampling either using standard MCMC, or by
increasing the size of the prior cache and continuing rejection sampling with
\thejoker.

@@

If just one posterior sample is returned after exhausting the full library of
prior samples, or if multiple (but fewer than \nposterior) are returned that all
lie within a single mode of the posterior \pdf, the posterior \pdf over orbital
parameters is treated as effectively unimodal: these stars are flagged ``needs
MCMC.''
In this case, we use the location of the returned sample (if only one is
returned), or a randomly chosen sample from those returned (if multiple samples
are returned within one mode) to generate a small Gaussian ball of initial
conditions and use standard MCMC to continue sampling until we obtain
\nposterior\ samples.

We use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010}) implemented
in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to perform the
samplings.
We transform the standard Keplerian orbital parameters to a safer
parametrization, $(\ln P, \sqrt{K}\,\cos M_0, \sqrt{K}\,\sin M_0, \sqrt{e}\,\cos
\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC sampling.
This reparametrization is safer and more efficient for sampling with
\package{emcee}, which expects parameters to be components of a vector so that
linear operations can be applied (see, e.g., \citealt{Hogg:2017}); the angle
variables $(\omega, M_0)$ don't meet this requirement in the standard
parametrization.
We use the same prior \pdf s as in \thejoker\ when running MCMC (see
\tablename~\ref{tbl:params}).

@@

If more than one posterior sample is returned after exhausting the full library
of prior samples, and the samples lie in multiple modes of the posterior \pdf,
the only way to proceed is to generate more prior samples and continue running
\thejoker: these stars are flagged ``needs more prior samples.''
In this case, we generate another equal-sized library of prior samples (a total
of $2\times\nprior$ samples) and re-do the rejection sampling.
We note that because of the way the rejection sampling step is done, this is not
equivalent to concatenating the results from a second, independent run of
\thejoker: the log-likelihood values for all of the prior samples must be used.
If at the end of this second run the target still has fewer than \nposterior\
samples, the sampling is flagged as ``incomplete.''

@@

We construct a comparison sample of simulated data with no RV variability to
assess our false-positive rates in the selections below (see
\sectionname~\ref{sec:catalogs}).
We randomly pick \ncontrol\ stars from the parent sample used in this work and
replace the visit velocity measurements with simulated data.
For each star $n$, we randomly sample an excess variance parameter value from
the prior, $s_n$.
For each visit $k$, we then sample a new velocity $v_{nk}$ by drawing from a
Gaussian with mean equal to the mean of the real data visit velocities,
$\bar{v}_n$, and variance equal to the sum of the visit variance (uncertainty),
$\sigma_{nk}$, and the excess variance,
\begin{equation}
    v_{nk} \sim \mathcal{N}(\bar{v}_n, \sigma_{nk}^2 + s_n^2) \quad .
\end{equation}
When we save the comparison data, we only store the visit uncertainty and
``forget'' the fact that the simulated data is generated with excess variance.

@@

\todo{Rix suggests this needs 1-2 more sentences to ease the transition: "It comes as an unexpected abrupt turn in the logical flow."}
As an initial use of the per-source posterior samplings, we use a hierarchical
Bayesian model to infer the parameters of the (assumed Gaussian) prior over the
log-excess-variance parameter, $(\mu_y, \sigma_y)$ (see
\tablename~\ref{tbl:params}).
This inference serves as a test-case for future work, where we intend to use the
independent posterior samplings to construct a hierarchical inference over
companion population properties.
This is also a test of the visit velocity uncertainties reported in the \apogee\
data products: If the catalog uncertainties are significantly underestimated, we
expect the inferred log-excess-variance distribution parameters to tend towards
larger values.

@@

We select a sample of stars that confidently have companions using percentiles
computed from the log of the posterior samples in the velocity semi-amplitude
parameter, $\ln K$.
\figurename~\ref{fig:lnK-percentiles}, shows the distribution of 1st percentiles
of $\ln K$ computed for all stars with posterior samplings from \thejoker\
(filled, blue histogram), along with the same for posterior samplings for the
null comparison sample (solid, dark line).
To select stars with companions, we use a cut in the 1st percentile of the $\ln
K$ samples such that $<1\%$ of the comparison sample is selected, i.e. our
estimated false positive rate is $<1\%$.
We use a threshold of $\ln K = 0$ to meet this constraint (vertical, dashed line
in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this cut.
For brevity below, we refer to this sample as the ``high-$K$'' stars, and the
complimentary sample as the ``low-$K$'' stars.

% Notebook: "Sample K cuts.ipynb"
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{lnK-percentiles}
\end{center}
\caption{%
Distribution of 1st percentiles in $\ln K$ (in units of \kms) for the parent
\apogee\ sample (blue, filled) and for the comparison sample (solid, dark line).
The vertical line (dashed, orange) indicates our adopted cut to select high-$K$
stars that likely have companions; $<1\%$ of the comparison sample falls above
this cut.
\label{fig:lnK-percentiles}
}
\end{figure}

\tablename~\ref{tbl:lnK-per} contains a list of all stars in the parent sample
and the value of the 1st percentile over the $\ln K$ samples for each star.
\figurename s~\ref{fig:highK-0} and \ref{fig:highK-1} show eight examples of
high-$K$ stars, i.e. stars that likely have companions, with different numbers
of visits, $N$, indicated on the panels.
\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} shows the same for eight
low-$K$ stars, i.e. stars that either have no companions, low-mass companions,
or companions at long periods or high inclination.
The majority of the stars in the high-$K$ sample have poorly constrained orbital
parameters because the posterior \pdf s are very multimodal.
The high-$K$ sample selected from \tablename~\ref{tbl:lnK-per} includes all
stars in the catalogs described in the next two subsections, however the
following two catalogs are mutually exclusive.

@@

Starting from the high-$K$ sample, we select stars that have posterior period
samples that fall within a single mode.
We define a period resolution for each star $\Delta =
\left[4\,P_{\textrm{min}}^2 / (2\pi \, T) \right]$, where $P_{\textrm{min}}$ is
the minimum period sample value, and $T$ is the epoch span of the data for that
star.
We consider a sampling to be unimodal when
\begin{equation}
    P_{\textrm{max}} - P_{\textrm{min}} < \Delta \label{eq:unimodal}
\end{equation}
(see \sectionname~2 of \citealt{Price-Whelan:2017}); \nunimodal\ stars
pass this cut.
Unlike the multi-modal stars, the posterior \pdf s for these stars can be
approximated using point-estimates and standard deviations of their respective
samplings.
We report maximum \textit{a posteriori} (MAP) sample values for the orbital
parameters, along with other computed quantities for all \nunimodal\ stars in
this high-$K$, unimodal sample.
For stars with measured primary masses, $M_1$, from other work
(\citealt{Ness:2015}), we compute the minimum companion mass, $M_{2,
\textrm{min}}$ and include both masses in this catalog.
We additionally join with the \apogee\ \DR\ \texttt{allStar} catalog and provide
all columns from this catalog for convenience.
\tablename~\ref{tbl:highK-unimodal} contains descriptions and units for all
columns in the high-$K$, unimodal sample catalog, available online.
\todo{Link to a webpage with the catalogs and info on how to use with
TwoBody/TheJoker}

We have visually inspected the inferred orbits for all systems in this sample
and have flagged systems with questionable or invalid fits.
The value of the flag is: 0, when the orbits look reasonable, 1, when the orbits
look reasonable but the value of the inferred excess variance is large, and 2,
when the orbits are clearly poor fits.
Many of the stars flagged as ``2'' look like they may be triple systems, as they
tend to have radial velocity variations over two distinct timescales that are
never well-fit by a single Keplerian orbit.
The stars flagged as ``1'' tend to be either upper RGB stars, where
astrophysical surface jitter can lead to RV modulations, or SB2 systems, where
absorption lines from the secondary confuse the RV pipeline and lead to strange
RV signals.
This flag is included in the catalog (\tablename~\ref{tbl:highK-unimodal}) as
the column \texttt{clean\_flag}; To select a clean sample of companions that
have been successfully vetted by-eye, select only stars with
\texttt{clean\_flag == 0}.

@@

A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
multimodal posterior distributions over orbital parameters that are limited to
a few qualitatively different solutions.
For these stars, one or a few more RV measurements would likely lead to
uniquely-determined orbital parameters; These stars are prime candidates for
follow-up efforts.
As examples of such cases, we include an additional catalog of systems that have
effectively bimodal posterior samplings in orbital period.
We identify these systems using $k$-means clustering (\citealt{Lloyd:1982}) of
the posterior samples in period.
In detail, for all stars in the high-$K$ sample, we compute $\ln P$ for all
period samples, then use $k$-means clustering with $k=2$ as implemented in the
\texttt{scikit-learn} package (\citealt{Pedregosa:2011}) to identify two
clusters of samples.
We initialize the cluster positions at either end of the range of possible
periods.
For each cluster, we then ask whether the samples assigned to that cluster are
unimodal (\eqname~\ref{eq:catalog-unimodal}).
If the samples in each respective cluster are unimodal, we call the sampling
``bimodal''; \nbimodal\ systems are identified as bimodal.

>>>>

\newcommand{\nprior}{536,870,912}
\newcommand{\nposterior}{256}
\newcommand{\nstars}{96,231}
\newcommand{\nvisits}{397,559}
\newcommand{\ncontrol}{16,384}
\newcommand{\lnKcut}{-0.2}
\newcommand{\nhighK}{4,898}
\newcommand{\nbimodal}{106}
\newcommand{\nunimodal}{322}

@@

\title{Binary companions of evolved stars in \apogee\ \DR: \\
       Search method and catalog of $\sim$5,000 companions}

@@

\begin{abstract}\noindent % trust me
% Context
Multi-epoch radial velocity measurements of stars can be used to identify
stellar, sub-stellar, and planetary-mass companions.
Even a small number of observation epochs can be useful for detecting
companions, though there can be multiple qualitatively different orbital
solutions that fit the data.
For these reasons, we custom-built a Monte Carlo sampler (\thejoker) that
delivers reliable (and often highly multi-modal) posterior samplings for
companion orbital parameters given sparse radial-velocity data.
% Aims
Here we perform a search for companions to \nstars\ red-giant stars observed in
the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
% Methods
We select stars with probable companions by making a cut on our posterior belief
about the amplitude of the stellar radial-velocity variation induced by the
orbit.
% Results
We provide (1)~a catalog of \nunimodal\ companions for which the stellar
companion properties can be confidently determined, (2)~a catalog of \nhighK\
stars that likely have companions, but for which more observations would be
needed to uniquely determine the orbital properties, and (3)~posterior samplings
for the full orbital parameters for all stars in the parent sample.
We highlight interesting systems and show the characteristics of systems with
confidently determined companion properties.
\end{abstract}

@@

Most of what is known about stellar companions comes from studies of
nearby main-sequence (MS) stars.
Nearly 50\% of MS stars in the solar neighborhood have companions
(e.g., \citealt{Tokovinin:2014}).
MS stars with companions have a large dynamic range of constituent and orbital
characteristics.
For example, binary stars have mass-ratios that span from $\approx 0.03$
to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
millions of years (e.g., \citealt{Raghavan:2010}).
Less is known about population properties of non-interacting or detached
companions to evolved stars.

@@

We have solved this problem previously with \thejoker\
(\citealt{Price-Whelan:2017}), a Monte Carlo rejection sampler that is
computationally expensive but probabilistically righteous:
It delivers independent posterior \pdf\ samples for single-companion binary
orbital parameters, given any number of radial-velocity measurements.
Here we use \thejoker\ to generate posterior \pdf\ samples for stars observed by
the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).

The \apogee\ surveys primarily target red-giant-branch (\RGB) and evolved stars,
which are ideal for the study of single-line binary systems.
In general, they are unlikely to have equally-bright companions, and their
spectra are therefore fit as single-line objects.
When this constraint is not met, \thejoker\ will in general fail, and a model
that fits for a mixture of stellar spectra is more appropriate (e.g.,
\citealt{El-Badry:2018}).
The subset of giant stars in the ``red clump'' (\RC) are even more powerful as
they are standard candles and have masses that can be estimated using
spectroscopy (using dredged-up elements; \citealt{Martig:2016,Ness:2016}).
With mass estimates for the primary star, the binary-orbit fitting will return
$M_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
of the so-called ``binary mass function.''
Additionally, the \apogee\ pipelines (\citealt{Garcia-Perez:2016}) and also
\thecannon\ (\citealt{Ness:2015}) produce detailed abundance estimates for \RGB\
and \RC\ stars.
If there are causal relationships between chemical abundances and binary
companions---as are expected---these should be measurable.

By making cuts on this library of posterior \pdf\ samples (described in detail
in \sectionname~\ref{sec:catalogs}), we deliver catalogs of binary-star systems
from the \apogee\ survey, and show the bulk properties of these systems.

@@

All data used in this work comes from the publicly-available data release 14
(\DR) of the \apogee\ survey (\citealt{Majewski:2017,Abolfathi:2017}), a
component of the Sloan Digital Sky Survey IV (\sdssiv;
\citealt{Gunn:2006,Blanton:2017}).
\apogee\ is designed to map stars across much of the Milky Way by obtaining
high-resolution ($R \sim 22,500$) infrared ($H$-band) spectroscopy of primarily
\RGB\ stars.
Targets are selected with simple color and brightness cuts, but the survey uses
fiber-plugged plates with a maximum of 300 fibers per each $\approx
1.5~\textrm{deg}^2$ field of view, leading to ``pencil-beam''-like sampling of
the Galactic stellar distribution.
In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
observed multiple times in a series of ``visits,'' typically with at least one
visit separated by a month or more in order to help identify binary stars.

@@

\thejoker\ is a custom-built Monte Carlo sampler designed to produce independent
posterior samples in Keplerian orbital parameters, given radial-velocity
measurements under the assumptions listed above.
Our parametrization of the orbital elements is similar to the notation in
\citet{Murray:2010}:
The radial velocity $v$ at time $t$ is given by
\begin{equation}
  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, M_0, t_0)\right) +
    e\,\cos\omega]
\end{equation}
where $\bs{\theta} = (P,e,M_0,\omega,K,v_0)$---period, eccentricity, mean
anomaly at a reference time $t_0$, argument of pericenter, velocity
semi-amplitude, barycentric velocity---and the true anomaly, $f$, is a function
of time and the specified parameters (see \sectionname~2 of
\citealt{Price-Whelan:2017} or \eqname~63 in \citealt{Murray:2010}).
In addition to the orbital parameters listed above, \thejoker\ can also generate
samples in an ``excess variance'' parameter, $s^2$, that is added to the
per-visit measurement variances.
This parameter allows us to test whether the visit RV uncertainties are
underestimated
For upper \RGB\ stars the inferred excess variance will be a combination of
extra systematic uncertainty and true astrophysical surface jitter
(e.g., \cite{Hekker:2008}).

\thejoker\ was designed for the extremely multi-modal \pdf 's expected when the
number of radial-velocity measurements of a source is small, or the data are
sparse (in phase-coverage) or noisy.
While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
independent samples with such data, \thejoker\ succeeds by brute force:
After generating an initial (very large) library of prior samples from an
assumed prior \pdf\ (see below), the (typically multi-modal) likelihood is
evaluated at each sample and used to rejection sample.
In practice, given a number of requested samples for each star, the sampling
proceeds iteratively: since it is easier to accept samples when the data is
sparse or noisy, far more prior sample draws (and thus likelihood evaluations)
must occur under very constraining data.

@@

For all \nstars\ \apogee\ stars with $\geq 3$ good visits (see
\sectionname~\ref{sec:data}), we use \thejoker\ to generate posterior samplings
for each star under the assumptions listed above (see
\sectionname~\ref{sec:fitting}).
We start by generating a library of \nprior\ prior samples generated under a
prior similar to that defined in \citet{Price-Whelan:2017}:
\begin{itemize}
    \item uniform or isotropic in angle parameters,
    \item uniform in log-period over the domain $[1,32768]~\textrm{day}$,
    \item a beta distribution over eccentricity (using parameters from
    \citealt{Kipping:2013}).
\end{itemize}
For the excess variance parameter, $s^2$, we use a Gaussian over the transformed
parameter $y = \ln s^2$ with the mean and standard deviation $(\mu_y, \sigma_y)$
indicated where the runs are described (see \sectionname~\ref{sec:catalogs}).
The reference time for each star is set to the minimum visit epoch; $M_0$ then
becomes the mean anomaly at the first visit observation.
\tablename~\ref{tbl:params} contains descriptions of all parameters and priors
used.

@@

We request \nposterior\ posterior samples for each source.
Depending on the data quality and phase coverage of the visits, \thejoker\ will
require different numbers of prior samples in order to rejection sample down to
the requested number of posterior samples: For few-epoch or noisy RV data, many
prior samples will pass the rejection step, whereas for very precise or
many-epoch RV data, \thejoker\ may need to process the full library of prior
samples.
We therefore generate the posterior samples using an iterative procedure that
adaptively predicts how many prior samples to test for each star.
For sources with very constraining data, \thejoker\ may return fewer than the
requested number of samples (as few as one sample).
When this occurs, we continue sampling either using standard MCMC, or by
increasing the size of the prior cache and continuing rejection sampling with
\thejoker.

@@

If just one posterior sample is returned after exhausting the full library of
prior samples, or if multiple (but fewer than \nposterior) are returned that all
lie within a single mode of the posterior \pdf, the posterior \pdf\ over orbital
parameters is treated as effectively unimodal: these stars are flagged ``needs
MCMC.''
In this case, we use the location of the returned sample (if only one is
returned), or a randomly chosen sample from those returned (if multiple samples
are returned within one mode) to generate a small Gaussian ball of initial
conditions and use standard MCMC to continue sampling until we obtain
\nposterior\ samples.

In detail, we use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010})
implemented in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to
perform the samplings.
We transform the standard Keplerian orbital parameters to a safer
parametrization, $(\ln P, \sqrt{K}\,\cos M_0, \sqrt{K}\,\sin M_0, \sqrt{e}\,\cos
\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC sampling.
This reparametrization is safer and more efficient for sampling with
\package{emcee}, which expects parameters to be components of a vector so that
linear operations can be applied (see, e.g., \citealt{Hogg:2017}); the angle
variables $(\omega, M_0)$ don't meet this requirement in the standard
parametrization.
We use the same prior \pdf 's as in \thejoker\ when running MCMC (see
\tablename~\ref{tbl:params}).

@@

If more than one posterior sample is returned after exhausting the full library
of prior samples, and the samples lie in multiple modes of the posterior \pdf,
the only way to proceed is to generate more prior samples and continue running
\thejoker: these stars are flagged ``needs more prior samples.''
In this case, we generate another equal-sized library of prior samples (a total
of $2\times\nprior$ samples) and re-do the rejection sampling.
We note that because of the way the rejection sampling step is done, this is not
equivalent to concatenating the results from a second, independent run of
\thejoker: the log-likelihood values for all of the prior samples must be used.
If at the end of this second run the target still has fewer than \nposterior\
samples, the sampling is flagged as ``incomplete.''
% TODO: this isn't done yet in the catalogs!

@@

We construct a comparison sample of simulated data with no RV variability to
assess our false-positive rates in the selections below (see
\sectionname~\ref{sec:catalogs}).
We randomly pick \ncontrol\ stars from the parent sample used in this work and
replace the visit velocity measurements with simulated data.
For each star $n$, we randomly sample an excess variance parameter value from
the prior, $s_n$.
For each visit $k$, we then sample a new velocity $v_{nk}$ by drawing from a
Gaussian with mean equal to the mean of the real data visit velocities,
$\bar{v}_n$, and variance equal to the sum of the visit variance (uncertainty),
$\sigma_{nk}$, and the excess variance,
\begin{equation}
    v_{nk} \sim \mathcal{N}(\bar{v}_n, \sigma_{nk}^2 + s_n^2) \quad .
\end{equation}
When we save the comparison data, we only store the visit uncertainty and
``forget'' the fact that the simulated data is generated with excess variance
(to be inferrred with \thejoker).

@@

% \todo{Rix suggests this needs 1-2 more sentences to ease the transition: "It comes as an unexpected abrupt turn in the logical flow."}
With posterior samplings for all \apogee\ \DR\ systems in hand, and as an
initial use of the per-source posterior samplings, we use a hierarchical
Bayesian model to infer the parameters of the (assumed Gaussian) prior over the
log-excess-variance parameter, $(\mu_y, \sigma_y)$ (see
\tablename~\ref{tbl:params}).
This inference serves as a test-case for future work, where we intend to use the
independent posterior samplings to construct a hierarchical inference over
companion population properties.
This is also a test of the visit velocity uncertainties reported in the \apogee\
data products: If the catalog uncertainties are significantly underestimated, we
expect the inferred log-excess-variance distribution parameters to tend towards
larger values.

@@

We select a sample of stars that confidently have companions using percentiles
computed from the log of the posterior samples in the velocity semi-amplitude
parameter, $\ln K$.
\figurename~\ref{fig:lnK-percentiles}, shows the distribution of 1st percentiles
of $\ln K$ computed for all stars with posterior samplings from \thejoker\
(filled, blue histogram), along with the same for posterior samplings for the
null comparison sample (solid, dark line).
To select stars with companions, we use a cut in the 1st percentile of the $\ln
K$ samples such that $1\%$ of the comparison sample is selected, i.e. our
estimated false positive rate is $\approx 1\%$.
We use a threshold of $\ln K = \lnKcut$ to meet this constraint (vertical,
dashed line in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this
cut.
For brevity below, we refer to this sample as the ``high-$K$'' stars, and the
complimentary sample as the ``low-$K$'' stars.

% Notebook: "Sample K cuts.ipynb"
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{lnK-percentiles}
\end{center}
\caption{%
Distribution of 1st percentiles in $\ln K$ (in units of \kms) for the parent
\apogee\ sample (blue, filled) and for the comparison sample (solid, dark line).
The vertical line (dashed, orange) indicates our adopted cut to select high-$K$
stars that likely have companions at $\ln K = \lnKcut$; $1\%$ of the comparison
sample falls above this cut.
\label{fig:lnK-percentiles}
}
\end{figure}

\tablename~\ref{tbl:lnK-per} contains a list of all stars in the parent sample
and the value of the 1st percentile over the $\ln K$ samples for each star.
\figurename s~\ref{fig:highK-0} and \ref{fig:highK-1} show eight examples of
high-$K$ stars, i.e. stars that likely have companions, with different numbers
of visits, $N$, indicated on the panels.
\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} shows the same for eight
low-$K$ stars, i.e. stars that either have no companions, low-mass companions,
or companions at long periods or high inclination.
The majority of the stars in the high-$K$ sample have poorly constrained orbital
parameters because the posterior \pdf 's are very multimodal.
The high-$K$ sample selected from \tablename~\ref{tbl:lnK-per} includes all
stars in the catalogs described in the next two subsections, however the
following two catalogs are mutually exclusive.

@@

Starting from the high-$K$ sample, we select stars that have posterior period
samples that fall within a single mode.
We define a period resolution for each star $\Delta =
\left[4\,P_{\textrm{min}}^2 / (2\pi \, T) \right]$, where $P_{\textrm{min}}$ is
the minimum period sample value, and $T$ is the epoch span of the data for that
star.
We consider a sampling to be unimodal when
\begin{equation}
    P_{\textrm{max}} - P_{\textrm{min}} < \Delta \label{eq:unimodal}
\end{equation}
(see \sectionname~2 of \citealt{Price-Whelan:2017}); \nunimodal\ stars
pass this cut.
Unlike the multi-modal stars, the posterior \pdf 's for these stars can be
approximated using point-estimates and standard deviations of their respective
samplings.
We report maximum \textit{a posteriori} (MAP) sample values for the orbital
parameters, along with other computed quantities for all \nunimodal\ stars in
this high-$K$, unimodal sample.
For stars with measured primary masses, $M_1$, from other work
(\citealt{Ness:2015}), we compute the minimum companion mass, $M_{2,
\textrm{min}}$ and include both masses in this catalog.
We additionally join with the \apogee\ \DR\ \texttt{allStar} catalog and provide
all columns from this catalog for convenience.
\tablename~\ref{tbl:highK-unimodal} contains descriptions and units for all
columns in the high-$K$, unimodal sample catalog, available online.
\todo{Link to a webpage with the catalogs and info on how to use with
TwoBody/TheJoker}

We have also visually inspected the inferred orbits for all systems in this
sample and have flagged systems with questionable or invalid fits.
The value of the flag is: 0, when the orbits look reasonable, 1, when the orbits
look reasonable but the value of the inferred excess variance is large, and 2,
when the orbits are clearly poor fits.
Many of the stars flagged as ``2'' look like they may be triple systems, as they
tend to have radial velocity variations over two distinct timescales that are
never well-fit by a single Keplerian orbit.
The stars flagged as ``1'' tend to be either upper RGB stars, where
astrophysical surface jitter can lead to RV modulations, or SB2 systems, where
absorption lines from the secondary confuse the RV pipeline and lead to strange
RV signals.
This flag is included in the catalog (\tablename~\ref{tbl:highK-unimodal}) as
the column \texttt{clean\_flag}; To select a clean sample of companions that
have been successfully vetted by-eye, select only stars with
\texttt{clean\_flag == 0}.

@@

A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
multimodal posterior distributions over orbital parameters that are limited to
a few qualitatively different solutions.
For these stars, one or a few more RV measurements would likely lead to
uniquely-determined orbital parameters; These stars are prime candidates for
follow-up efforts.
As examples of such cases, we include an additional catalog of systems that have
effectively bimodal posterior samplings in orbital period.
We identify these systems using $k$-means clustering (\citealt{Lloyd:1982}) of
the posterior samples in period.
In detail, for all stars in the high-$K$ sample, we compute $\ln P$ for all
period samples, then use $k$-means clustering with $k=2$ as implemented in the
\texttt{scikit-learn} package (\citealt{Pedregosa:2011}) to identify two
clusters of samples.
We initialize the cluster positions at either end of the range of possible
periods.
For each cluster, we then ask whether the samples assigned to that cluster are
unimodal (\eqname~\ref{eq:unimodal}).
If the samples in each respective cluster are unimodal, we call the sampling
``bimodal''; \nbimodal\ systems are identified as bimodal.
